Classless Java: Tuning Java Interfaces

### 2015-12-07 paper/sections/FutureWork.tex
Added new version of supporting privateness

\subsection{Qualifiers in Methods} %{Private state
The biggest limitation of our approach is the inability to model qualifiers
for class methods (private, protected, synchronized, etc.). For example, the absence
of support for private/protected methods in Java 8 interfaces forces all
members of interfaces to be public, including static methods. Since we use
abstract methods to encode state, our state is always all public. Still, because 
the state can only be accessed by methods, it 
is impossible for the user to know if a certain method maps directly to a field
or if it has a default implementation.  If the user wants a constructor that
does not directly maps to the fields, (as for secondary constructors in Scala)
he can simply define its own \Q@of@ method and delegate on the generated one:
\begin{lstlisting}
@Obj interface Point{  int x(); int y();
  static Point of(int val){return Point.of(val,val);}  }
\end{lstlisting}
However, the generated \Q@of@ method would also be present and public.  If a
future version of Java was to support \emph{static private methods in
  interfaces} we could extend our code generation to handle also encapsulation.
Currently, it is possible to use a public nested class with private static
methods inside, but this is ugly and cumbersome. One possibility is that the
annotation processor takes methods with a \Q!@Private! annotation, and turns it into
static private methods of a nested class. In this extension, also the \Q@of@
method could be made private following the same pattern.

% \yanlin{future work on state initialization removed for now. Because I think the
%   other two future works are fine. But this one is still not clear and
%   controversial. The proposed workaround forces users to make decision: if he
%   wants state with initialization then he also has to define the setter. Or
%   else, it is ambiguous: could be interpreted as a state initializer or just
%   a normal default method.}
\begin{comment}
\subsection{State initialization}
As discussed before, the user can trivially define its own \Q@of@ method, and
initialize a portion of the state with default values.  However, the
initialization code would not be reusable, and subinterfaces would have
to repeat such initialization code.  If a field has no setters, a simple
alternative is to just define the ``field'' as a default method as in
\begin{lstlisting}
@Obj interface Box{ default int val(){return 0;} }
\end{lstlisting}
If setters are required, a possible extension of our code expansion could
recognize a field if the getter is provided and the setter is required, and
could generate the following code:
\lstinputlisting[linerange=4-11]{../UseMixinLombok/src/test/TestStateInitialization.java}% APPLY:linerange=STATE_INIT

>>>>

\subsection{Encapsulation}
The biggest limitation of our approach is the inability to model visibility restrictions. For example, the absence
of support for private/protected methods in Java 8 interfaces forces all
members of interfaces to be public, including static methods. Since we use
abstract methods to encode state, our state is always all public. Still, because 
the state can only be accessed by methods, it 
is impossible for the user to know if a certain method maps directly to a field
or if it has a default implementation.  If the user wants a constructor that
does not directly maps to the fields, (as for secondary constructors in Scala)
he can simply define its own \Q@of@ method and delegate on the generated one:
\begin{lstlisting}
@Obj interface Point{  int x(); int y();
  static Point of(int val){return Point.of(val,val);}  }
\end{lstlisting}
However, the generated \Q@of@ method would also be present and public.  If a
future version of Java was to support \emph{static private methods in
  interfaces} we could extend our code generation to handle also encapsulation.

However, interfaces as a whole can have public or 
package private (java default) visiblity.

We can add a second annotation 
\Q!@Exposed!
that leverage on this edge:
An
interface without exposed works as usual, but if any method of a public \mixin interface is annoted with \Q!@Exposed!, we can apply a translation where a new  (package private) interface type is introduced, and the orginal interface become just a facade.
For example:
\begin{lstlisting}
@Obj public interface Person{
  void name(String val);
  @Exposed default void rename(String newName){ if(/*valid name*/){ this.name(val);}}
  @Exposed String name();
  @Exposed static Person from(String val){ if(/*valid name*/){return Person.of(val);}
    throw /*invalid name*/}  }
\end{lstlisting}
becomes
\begin{lstlisting}
public interface Person{
  void rename(String newName)
  String name();
  static Person from(String val){ return Person$.from(val);} }

@Obj interface Person$ extends Person{//will be further expanded by @Obj
  void name(String val);
  default void rename(String newName){ if(/*valid name*/){ this.name(val);}}
  String name();
  static Person from(String val){ if(/*valid name*/){return Person$.of(val);}
    throw /*invalid name*/}  }
\end{lstlisting}

 This is not a perfect solution, since \Q@Person$@ can still be seen inside the \Q@Person@ package and heirs of \Q@Person$@, however it is surprising we achive such of a good result without any language support for privatness in interfaces.

\begin{comment}
\subsection{Qualifiers in Methods} %{Private state
The biggest limitation of our approach is the inability to model qualifiers
for class methods (private, protected, synchronized, etc.). For example, the absence
of support for private/protected methods in Java 8 interfaces forces all
members of interfaces to be public, including static methods. Since we use
abstract methods to encode state, our state is always all public. Still, because 
the state can only be accessed by methods, it 
is impossible for the user to know if a certain method maps directly to a field
or if it has a default implementation.  If the user wants a constructor that
does not directly maps to the fields, (as for secondary constructors in Scala)
he can simply define its own \Q@of@ method and delegate on the generated one:
\begin{lstlisting}
@Obj interface Point{  int x(); int y();
  static Point of(int val){return Point.of(val,val);}  }
\end{lstlisting}
However, the generated \Q@of@ method would also be present and public.  If a
future version of Java was to support \emph{static private methods in
  interfaces} we could extend our code generation to handle also encapsulation.
Currently, it is possible to use a public nested class with private static
methods inside, but this is ugly and cumbersome. One possibility is that the
annotation processor takes methods with a \Q!@Private! annotation, and turns it into
static private methods of a nested class. In this extension, also the \Q@of@
method could be made private following the same pattern.
\end{comment}
\marco{old version still commented if we decided to swap back}
% \yanlin{future work on state initialization removed for now. Because I think the
%   other two future works are fine. But this one is still not clear and
%   controversial. The proposed workaround forces users to make decision: if he
%   wants state with initialization then he also has to define the setter. Or
%   else, it is ambiguous: could be interpreted as a state initializer or just
%   a normal default method.}
\begin{comment}
\subsection{State initialization}
As discussed before, the user can trivially define its own \Q@of@ method, and
initialize a portion of the state with default values.  However, the
initialization code would not be reusable, and subinterfaces would have
to repeat such initialization code.  If a field has no setters, a simple
alternative is to just define the ``field'' as a default method as in
\begin{lstlisting}
@Obj interface Box{ default int val(){return 0;} }
\end{lstlisting}
If setters are required, a possible extension of our code expansion could
recognize a field if the getter is provided and the setter is required, and
could generate the following code:
\lstinputlisting[linerange=4-11]{../UseMixinLombok/src/test/TestStateInitialization.java}% APPLY:linerange=STATE_INIT
